### 2024-11-25 code/01_data_checks.qmd
edit manuscript

---
title: "Data checks"
author: "Stefano Coretta"
format: html
---

```{r}
#| label: setup

library(tidyverse)
library(coretta2018itaegg)
library(Hmisc)
```

```{r}
#| label: load-data
#| 
data("ita_egg")
```

```{r}
ita_egg |> 
  ggplot(aes(voi_clo_prop)) +
  geom_density()

ita_egg |> 
  filter(voi_clo_prop > 0, voi_clo_prop < 1) |> 
  ggplot(aes(voi_clo_prop)) +
  geom_density() +
  geom_rug()
```

```{r}
ita_egg_filt <- ita_egg |> 
  filter(voi_clo_prop > 0, voi_clo_prop < 1)
```

```{r}
ita_egg_filt |> 
  ggplot(aes(vowel, voi_clo_prop)) +
  geom_jitter(alpha = 0.1, width = 0.2) +
  stat_summary(fun.data = "mean_cl_boot")
```

```{r}
ita_egg_filt |> 
  ggplot(aes(vowel, voi_clo_prop)) +
  geom_jitter(alpha = 0.2, width = 0.2, aes(colour = vowel)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(cols = vars(c2))
```

```{r}
ita_egg_filt |> 
  ggplot(aes(vowel, voi_clo_prop)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(cols = vars(c2))
```

```{r}
ita_egg_filt |> 
  ggplot(aes(speech_rate, voi_clo_prop)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  facet_wrap(vars(vowel))
```