### 2025-02-06 manuscript/manuscript.tex
edit manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}

\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A tutorial of Bayesian beta regressions with brms in R},
  pdfauthor={Stefano Coretta; Paul Bürkner},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{A tutorial of Bayesian beta regressions with brms in R}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk} 
\author[]{Paul Bürkner%
%
}
 \ead{paul.buerkner@gmail.com}

\affiliation[1]{organization={University of Edinburgh, Linguistics and
English Language},addressline={3 George
Sq},city={Edinburgh},postcode={EH8 9AD},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{introduction}

Phonetic research often involves numeric continuous outcome variables,
like durations, frequencies, loudness and ratios
\citep{kluender1988, johnson2003, gick2013, ladefoged2014, abramson2017, roettger2018, coretta2023}.
Another commonly employed type of outcome variable are proportions: for
example, proportion of voicing during closure \citep{davidson2016},
vocal folds contact quotient \citep{herbst2017}, gesture amplitude
\citep{carignan2021}, nasalance \citep{carignan2021}. Moreover,
virtually any measure can be MIN-MAX normalised, a procedure which
transforms values so that they are in the range 0--1.

Regression models (and extensions like generalised additive models) have
become a \emph{de facto} standard for the statistical analysis of a
variety of measures in phonetic research
\citep{kirby2018, politzerahles2018, tavakoli2024}. However, there is a
tendency for researchers to use Gaussian distribution families
(i.e.~probability distributions for the outcome variable) for any
measure that is numeric, irrespective of whether the measure is
unbounded, like in truly Gaussian variables, or bounded, like in
proportions. A possible reason is that the base R function for fitting
regression models, \texttt{lm()}, and the lme4 function used to fit
regression models with varying terms, \texttt{lmer()} from lme4
\citep{bates2015}, both fit Gaussian regressions by default and the user
does not have to specify the distribution family. This tacit default of
using Gaussian models is also reflected in teaching practices, where
significance test and models using the Gaussian distribution are the
first to be taught \citep{baayen2008, winter2020}, due to their relative
simplicity and the fact that regression models with other families are
conceptual generalisations of Gaussian regression models.

While most researchers approach proportions with Gaussian regression
models, proportion are not Gaussian by nature, since they are continuous
variables bounded between 0 and 1. Thus, regression models with
proportions as outcome variable should be fitted using a likelihood
family that assumes such unit interval data. A common and flexible
distribution for this purpose is the beta distribution. This tutorial
introduces researchers to beta regression models in R using the package
brms. Familiarity with regression modelling in R with a package like
lme4 is assumed, but no prior knowledge of Bayesian statistics is
necessary. The rest of the paper is structured as follows:
Section~\ref{sec-beta} introduces the mathematical underpinnings of the
beta distribution, Section~\ref{sec-bayes} provides the readers with a
brief explanation and justification of Bayesian regression models, while
Section~\ref{sec-case-1} and Section~\ref{sec-case-2} illustrate how to
fit Bayesian beta regression models with two case studies.

\section{The beta distribution}\label{sec-beta}

The beta distribution is often considered as \emph{the} distribution for
the modelling of unit interval data
\citep{ferrari_beta_2004, cribari-neto_beta_2010}. It has been
extensively studied theoretically
\citep{krysicki_new_1999, gupta_handbook_2014, espinheira_beta_2008} and
is used a baseline to compare other unit interval distributions against
\citep{kieschnick_regression_2003, bonat_regression_2013, lopez_bayesian_2013}.
Specifically in Bayesian regression models, it was shown to perform well
across a wide range of scenarios
\citep{scholz_posterior_2023, scholz_prediction_2025}, which is why we
focus on the beta distribution in this tutorial.

We use a common mean parametrisation for the beta distribution with mean
parameter \(\mu\), bounded between 0 and 1, and positive precision
parameter \(\phi\) that is roughly proportional to the inverse of the
variance \(\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}\). That is,
the larger \(\phi\) the smaller the variance of the corresponding beta
distribution. While the mathematical details are not needed to
understand the content of this tutorial, we still write down the density
below for reasons of completeness:

\[
p(y \mid \mu, \phi) = \frac{1}{B(\mu \, \phi, \; (1-\mu) \, \phi)} \; y^{\mu \, \phi -1} \; (1-y)^{(1-\mu) \, \phi -1},
\]

where \(B\) is the Beta function, a complex integral for whose numerical
approximation efficient algorithms exist in every programming language.

\section{Bayesian regression models}\label{sec-bayes}

Bayesian regression models are being increasingly adopted within
phonetics and language research more broadly
\citep{vasishth2018, nalborczyk2019, verissimo2021}. Bayesian inference
involves updating of prior probability distributions in light of
evidence from data, to produce posterior probability distributions. In
Bayesian regression models, model parameters are modelled as full
probability distributions, rather than point estimates as in Null
Hypothesis Significance Testing. Given the difficulty of analytical
solutions of model equations, Bayesian regressions rely on sampling
algorithms to reconstruct the posterior distributions. The statistical
language Stan \citep{standevelopmentteam2017} employs efficient and
robust Markov Chain Monte Carlo algorithms for fitting a variety of
models, and the R package brms allows R users to interface with Stan
from within R to fit Bayesian regression models
\citep{burkner2017, burkner2018, burkner2021}.

The main practical advantage of Bayesian regression models over
maximum-likelihood-based frequentist regression models, like those
fitted with the lme4 package \citep{bates2015}, is that Bayesian
regression models don't suffer from the convergence issues that models
fitted in lme4 \citep{bates2015} and other packages
\citep{cribarineto2010} that fit frequentist models do, independent of
sample size. A second, long-term advantage is that Bayesian regression
models allow researchers to statistically re-use information from
previous studies by specifying informative priors. While prior
specification is one of the main features of Bayesian inference, in this
tutorial we will use the default priors as set by brms. These are
sensible priors estimated from the data that facilitate convergence but
bear virtually no influence on the estimated posteriors. Specifying
priors requires a great deal of precise quantitative knowledge, which in
most areas of phonetics we still do not possess, so that using default
uninformative priors is, for the time being, theoretically sound.

The output of Bayesian regression models is (posterior) probability
distributions for the model parameters, through which researchers can
quantify (un)certainty. Bayesian Credible Intervals (CrIs) can be
calculated from the posterior distributions at several probability
levels (e.g.~95, 90, 80, 60\%) for a more complete view on estimated
parameters. Bayesian CrIs indicate that at a certain probability levels
the parameter's estimate lies within that interval: so, for example, a
90\% CrI {[}A, B{]} indicates that there is a 90\% probability that the
estimate is between A and B. Different probability levels correspond to
different levels of confidence: the higher the probability the higher
the confidence (always conditional on data and model). Readers
interested in a full and accessible exposition of Bayesian statistics
are referred to \citet{mcelreath2019}. Shorter introductions can be
found in \citet{etz2018}, \citet{vasishth2018} and
\citet{nalborczyk2019}.

\section{Case study 1: voicing within consonant
closure}\label{sec-case-1}

For the first case study, we will model the proportion of voicing within
consonant closure. The measurements come from a data set of audio and
electroglottographic (EGG) recordings of 19 speakers of Northwestern
Italian \citep{coretta2019k, coretta2020b}. The participants read frame
sentences which included target words of the form /CVCo/, where /C/ was
either /k, t, p/ in all permutations and /V/ was either /i, e, a, ɔ, u/
(two resulting words, /peto/ and /kako/ were excluded because they are
profanities), for a total of 43 target words. There were 4 different
frame sentence: \emph{Scrivete X sul foglio} `Write X on the sheet',
\emph{Ha detto X sei volte} `She said X six times', \emph{Sentivo X di
nuovo} `I heard X again', \emph{Ripete X da sempre} `She's been
repeating X since ever'. There is a total of 172 trials per participant
(3,268 grand total). The actual observation count is 2,419, after
removing speech errors, EGG measurement errors, and cases in which
voicing ceased before the closure onset/after the closure offset of the
second /C/.

The proportion of voicing during the closure of the second /C/ was
calculated as the proportion of contiguous voicing duration after
closure onset to total duration of closure. The following code chunk
attaches the tidyverse packages \citep[for reading and wrangling
data,][]{wickham2019} and loads the \texttt{ita\_egg} tibble (data
frame). The tibble is filtered so as to remove voicing proportions
(\texttt{voi\_clo\_prop}) that are smaller than 0 and greater than 1.
The variables \texttt{vowel} and \texttt{c2} are converted to factors to
specify the order of the levels.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach tidyverse and set light theme for plots}
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{theme\_set}\NormalTok{(}\FunctionTok{theme\_light}\NormalTok{())}

\CommentTok{\# load tibble}
\FunctionTok{load}\NormalTok{(}\StringTok{"data/coretta2018/ita\_egg.rda"}\NormalTok{)}

\CommentTok{\# filter and mutate data}
\NormalTok{ita\_egg }\OtherTok{\textless{}{-}}\NormalTok{ ita\_egg }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{filter}\NormalTok{(voi\_clo\_prop }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{, voi\_clo\_prop }\SpecialCharTok{\textless{}} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{vowel =} \FunctionTok{factor}\NormalTok{(vowel, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"i"}\NormalTok{, }\StringTok{"e"}\NormalTok{, }\StringTok{"a"}\NormalTok{, }\StringTok{"o"}\NormalTok{, }\StringTok{"u"}\NormalTok{)),}
    \AttributeTok{c2 =} \FunctionTok{factor}\NormalTok{(c2, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"k"}\NormalTok{, }\StringTok{"t"}\NormalTok{, }\StringTok{"p"}\NormalTok{))}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

Table~\ref{tbl-ita-egg} shows the first ten rows of the tibble (only
relevant columns are included).

\begin{longtable}[]{@{}llllr@{}}

\caption{\label{tbl-ita-egg}First 10 rows of the data of voicing
proportion within stop closure.}

\tabularnewline

\toprule\noalign{}
speaker & word & vowel & c2 & voi\_clo\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
it01 & poto & o & t & 0.0457950 \\
it01 & topo & o & p & 0.3522513 \\
it01 & pato & a & t & 0.1440749 \\
it01 & teto & e & t & 0.3610916 \\
it01 & toto & o & t & 0.2743519 \\
it01 & puco & u & k & 0.1943984 \\
it01 & chipo & i & p & 0.2227896 \\
it01 & peco & e & k & 0.1835596 \\
it01 & poco & o & k & 0.1771007 \\
it01 & poto & o & t & 0.1525163 \\

\end{longtable}

Figure~\ref{fig-ita-egg} shows the raw voicing duration proportion
values split by vowel /i, e, a, ɔ, u/ and second consonant /k, t, p/ in
the /CVCo/ target words. The plot suggests that, on average, the voicing
proportion is slightly lower with /k/ than with /t, p/. Moreover, there
is greater variability between vowels in /t, p/ than in /k/. We will use
a beta regression model to assess these patterns. {[}``expectations''
XXX{]}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-ita-egg-1.pdf}}

}

\caption{\label{fig-ita-egg}Proportion of voicing during the closure of
the second consonant in /CVCo/ words by vowel and the second consonant.}

\end{figure}%

We will use brms to fit Bayesian beta regressions \citep{burkner2017}.
The model has voicing proportion as the outcome variable and the
following terms: an interaction between vowel (/i, e, a, ɔ, u/) and
second consonant C2 (/k, t, p/), centred speech rate (number of
syllables per second); as varying (aka random) terms, by-speaker varying
coefficients for the vowel/consonant interaction and for centred speech
rate.\footnote{Footnote about Gelman's terminology for random effects.}
The categorical predictors vowel and C2 are coded using indexing rather
than traditional R contrasts: in R, this corresponds to suppressing the
model's intercept with the \texttt{0\ +} syntax; using indexing instead
of contrasts makes it easier to specify priors. For pedagogical
simplicity, the model will use the default priors, but note that in real
data analyses contexts, priors should be specified by the user. I refer
the readers to XXX.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach brms}
\FunctionTok{library}\NormalTok{(brms)}

\CommentTok{\# fit the model}
\CommentTok{\# Takes 3 minutes on MacBook Pro 2020, M1}
\NormalTok{voi\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
  \CommentTok{\# model formula}
\NormalTok{  voi\_clo\_prop }\SpecialCharTok{\textasciitilde{}}
    \CommentTok{\# constant terms}
    \DecValTok{0} \SpecialCharTok{+}\NormalTok{ vowel}\SpecialCharTok{:}\NormalTok{c2 }\SpecialCharTok{+}\NormalTok{ speech\_rate\_c }\SpecialCharTok{+}
    \CommentTok{\# varying terms}
\NormalTok{    (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ vowel}\SpecialCharTok{:}\NormalTok{c2 }\SpecialCharTok{+}\NormalTok{ speech\_rate\_c }\SpecialCharTok{|}\NormalTok{ speaker),}
  \CommentTok{\# uses the beta family for the outcome}
  \AttributeTok{family =}\NormalTok{ Beta,}
  \AttributeTok{data =}\NormalTok{ ita\_egg,}
  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
  \AttributeTok{file =} \StringTok{"data/cache/voi\_prop\_bm"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
full summary with an explanation of each part can be found in XXX.
Table~\ref{tbl-voi-prop-fixef} reports the output of \texttt{fixef()} as
a table (we round all values to the nearest 2 digits for clarity.). For
each coefficient in the model, \texttt{fixef()} prints the name of the
coefficient, the mean estimate, the estimate error and the lower and
upper limits of a Bayesian Credible interval (CrI). Here, we print an
80\% CrI. There is nothing special about 95\% CrI within Bayesian
inference and instead experts recommend to check and report a variety of
CrIs. Obtaining CrIs at different probability levels allows researchers
to make more fine-grained inferential statements than the frequentist
significance dichotomy affords. For simplicity of exposition, we will
use 80\% CrIs in this case study but we strongly recommend researchers
to always obtain CrIs at different levels of probability and base their
inferences on all and not one in particular. To reiterate, in Bayesian
inference, an 80\% CrI indicates the range of values within which the
true estimate falls at 80\% probability or confidence.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-voi-prop-fixef}Regression coefficients of a beta
regression of voicing proportion (\texttt{voi\_prop\_bm}).}

\tabularnewline

\toprule\noalign{}
& Estimate & Est.Error & Q10 & Q90 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
speech\_rate\_c & 0.08 & 0.06 & 0.01 & 0.15 \\
voweli:c2k & -0.91 & 0.14 & -1.08 & -0.74 \\
vowele:c2k & -1.08 & 0.11 & -1.22 & -0.94 \\
vowela:c2k & -0.99 & 0.12 & -1.14 & -0.84 \\
vowelo:c2k & -0.79 & 0.14 & -0.96 & -0.62 \\
vowelu:c2k & -1.00 & 0.16 & -1.20 & -0.80 \\
voweli:c2t & -0.66 & 0.11 & -0.79 & -0.53 \\
vowele:c2t & -0.84 & 0.14 & -1.02 & -0.66 \\
vowela:c2t & -1.43 & 0.13 & -1.60 & -1.26 \\
vowelo:c2t & -1.15 & 0.13 & -1.31 & -0.99 \\
vowelu:c2t & -0.68 & 0.12 & -0.83 & -0.54 \\
voweli:c2p & -0.68 & 0.11 & -0.81 & -0.54 \\
vowele:c2p & -0.88 & 0.15 & -1.07 & -0.68 \\
vowela:c2p & -1.14 & 0.13 & -1.31 & -0.98 \\
vowelo:c2p & -0.66 & 0.11 & -0.80 & -0.53 \\
vowelu:c2p & -0.44 & 0.12 & -0.59 & -0.28 \\

\end{longtable}

The coefficients of a beta regression are estimated on the log-odds
scale, as in Bernoulli/binomial (aka logistic) regressions. From the
summary, we see that speech rate (number of syllables per second) has a
positive effect on voicing proportion: the 80\% CrI is between 0.01 and
0.15 log-odds {[}\(\beta\) = 0.08, SD = 0.06{]}. Log-odds can be
converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds
correspond to an odd-ratio of 1.01 to 1.16, or as percentages, to an
increase of voicing of 1 to 16\% for every increase of one syllable per
second. Since this is an 80\% CrI, we can be 80\% confident that the
true effect of speech rate is between 1-16\% increase of voicing
proportion, conditional on the data and model. Note that transforming
measures this way is appropriate \emph{only} with quantile-based
measures (like CrIs) but not with moments like the mean and standard
deviation: to correctly get mean and SDs in the transformed scale, you
must first extract the posterior draws (see below), convert them and
then take moments such as mean and SD (for a more detailed explanation,
see XXX). In the avoidance of doubt, we will always transform the drawn
values first and then take summary measures.

Turning now to the coefficients for vowel and C2, given the indexing
approach of coding these variables the model summary and the output of
\texttt{fixef()} reports the \emph{predictions} in log-odds for each
combination of vowel and C2, rather than differences between levels. The
CrIs of the vowel/C2 coefficients span all negative log-odds values:
these correspond to proportions that are lower than 0.5 (which is 0 in
log-odds). This matches the general trends in the raw data, which we
plotted in Figure~\ref{fig-ita-egg}.

Next, we will plot the predicted proportions of each vowel/C2
combination at mean speech rate (i.e.~centred speech rate = 0) and then
calculate the average pair-wise difference in voicing proportion between
/k, t, p/. Finally, we will assess whether there is greater
between-vowel variability in /t, p/ relative to /k/.

Before being able to plot the predictions, it's important to get
familiar with the so-called posterior draws. {[}Bayesian MCMC XXX{]}.
Posterior draws can be conveniently obtained with the
\texttt{as\_draws\_df()}. For the moment, we will extract only the draws
of the constant regression coefficients (model variables starting with
\texttt{b\_}). To check which coefficients are available in a model, use
\texttt{get\_variables()} from the tidybayes package.
\texttt{as\_draws\_df()} returns a tibble where each column contains the
drawn values of a coefficient. Table~\ref{tbl-voi-prop-draws} shows the
first ten rows and first five columns of the output of
\texttt{as\_draws\_df()}. The probability distribution of the drawn
values of each coefficient is the posterior probability distribution of
that coefficient. Note that, due to the indexing coding of vowel and C2,
all coefficient except \texttt{b\_speech\_rate\_c} are \emph{predicted
log-odds} for each vowel/C2 combination (the drawn values for
\texttt{b\_speech\_rate\_c} are drawn \emph{differences} in log-odds for
each unit increase of speech rate). The drawn values are in log-odds,
but we can convert them to proportions with \texttt{plogis()} (we will
do this when plotting below).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{voi\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: Dropping 'draws_df' class as required metadata was removed.
\end{verbatim}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2353}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}@{}}

\caption{\label{tbl-voi-prop-draws}First ten rows and 5 columns of the
posterior draws for the model \texttt{voi\_prop\_bm}.}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
b\_speech\_rate\_c
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_voweli:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowele:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowela:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowelo:c2k
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.1593890 & -0.7213516 & -1.046577 & -1.1197185 & -1.0191525 \\
0.0351033 & -0.7492306 & -1.385450 & -1.0633164 & -0.8854320 \\
0.1259352 & -0.8472495 & -0.975776 & -1.0972735 & -0.6440836 \\
0.0640492 & -0.7033469 & -1.068626 & -1.0475190 & -0.8328252 \\
0.0319081 & -0.7659281 & -1.177656 & -0.9357499 & -0.6955130 \\
0.0729965 & -0.8494392 & -1.226692 & -1.0635364 & -0.7692243 \\
0.1009609 & -0.9546148 & -1.072955 & -1.1408255 & -0.8566750 \\
0.0735392 & -1.0160675 & -1.065025 & -1.1736263 & -0.8334884 \\
0.0870061 & -1.0436475 & -1.069337 & -1.2632061 & -0.7941857 \\
0.0923815 & -1.2042168 & -1.041238 & -0.9158689 & -0.6444492 \\

\end{longtable}

We can now wrangle this tibble and plot the posterior distributions for
each vowel/C2 combination. Table~\ref{tbl-voi-long} shows the first ten
rows.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# drop b\_speech\_rate\_c before pivoting}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{b\_speech\_rate\_c) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot vowel:c2 columns}
  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voweli:c2k}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_vowelu:c2p}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{, }\StringTok{"c2"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrlllr@{}}

\caption{\label{tbl-voi-long}First ten rows of posterior draws from
\texttt{voi\_prop\_bm} in long format.}

\tabularnewline

\toprule\noalign{}
.chain & .iteration & .draw & type & vowel & c2 & value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 1 & 1 & b & voweli & c2k & -0.7213516 \\
1 & 1 & 1 & b & vowele & c2k & -1.0465766 \\
1 & 1 & 1 & b & vowela & c2k & -1.1197185 \\
1 & 1 & 1 & b & vowelo & c2k & -1.0191525 \\
1 & 1 & 1 & b & vowelu & c2k & -1.0523007 \\
1 & 1 & 1 & b & voweli & c2t & -0.6248875 \\
1 & 1 & 1 & b & vowele & c2t & -0.7951552 \\
1 & 1 & 1 & b & vowela & c2t & -1.3999773 \\
1 & 1 & 1 & b & vowelo & c2t & -1.0950180 \\
1 & 1 & 1 & b & vowelu & c2t & -0.6682030 \\

\end{longtable}

For plotting, we can use ggplot2 statistics layers from the ggdist
package. \texttt{stat\_halfeye()} plots the density of the posterior
probability (in grey), its median (point) and CrIs (lines). Let's use 60
and 80\% CrIs and transform the log-odds values to proportions with
\texttt{plogis()}. See Figure~\ref{fig-voi-prop-bm-1}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach ggdist package}
\FunctionTok{library}\NormalTok{(ggdist)}

\NormalTok{voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value), vowel)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-1-1.pdf}}

}

\caption{\label{fig-voi-prop-bm-1}Expected predictions of voicing
proportion by vowel and consonant from \texttt{voi\_prop\_bm}.}

\end{figure}%

What if we want to plot the average predicted voicing proportion for the
three consonants /k, t, p/? One approach is to take the mean across
vowels within each consonant for each posterior draw, and the posterior
distribution of the resulting list of values is the predicted posterior
distribution of voicing proportion for each consonant, assuming an
``average'' vowel.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and c2 ensures that averaging applies only within draw and c2}
  \FunctionTok{group\_by}\NormalTok{(.draw, c2) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# calculate the mean value within draw/c2}
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), c2)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-pbdl-c2-1.pdf}}

}

\caption{\label{fig-voi-pbdl-c2}Expected predictions of voicing
proportion by consonant, averaged across vowels.}

\end{figure}%

Based on the expected posterior distributions of the mean voicing
proportion by consonant, /p/ has a somewhat higher voicing proportion
than /k/ and /t/. The real question is: how much higher? We can quantify
this by taking the difference of the drawn values for /p/ and those for
/t, k/ (all averaged across vowels). Since we want to compare /t, k/
with /p/, we should first average the average draws of /t, k/ and then
take the difference of the averaged /t, k/ draws and the draws of /p/.
Table~\ref{tbl-voi-pbd} shows the first ten rows of the resulting data
frame. The posterior distribution of the expected difference is shown in
Figure~\ref{fig-diff-p-tk}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per consonant with the mean drawn values,}
  \CommentTok{\# with one draw per raw}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ c2, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the mean of /k/ and /t/, for each draw}
    \AttributeTok{c2tk =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{c}\NormalTok{(c2k, c2t)),}
    \CommentTok{\# calculate the difference of /p/ and /t, k/}
    \AttributeTok{c2p\_tk\_diff =}\NormalTok{ c2p }\SpecialCharTok{{-}}\NormalTok{ c2tk}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrrrr@{}}

\caption{\label{tbl-voi-pbd}First 10 rows of expected difference of
voicing proportion between /t, k/ and /p/.}

\tabularnewline

\toprule\noalign{}
.draw & c2k & c2p & c2t & c2tk & c2p\_tk\_diff \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & -0.9918200 & -0.7450832 & -0.9166482 & -0.9533426 & 0.2082595 \\
2 & -1.0240673 & -0.7904348 & -1.0322325 & -0.9533426 & 0.1629079 \\
3 & -0.8954946 & -0.6873228 & -0.8904412 & -0.9533426 & 0.2660199 \\
4 & -0.9442361 & -0.8135191 & -0.9095122 & -0.9533426 & 0.1398235 \\
5 & -0.9278262 & -0.7908457 & -0.9200903 & -0.9533426 & 0.1624970 \\
6 & -0.9825895 & -0.7283054 & -0.9262196 & -0.9533426 & 0.2250372 \\
7 & -1.0482359 & -0.8731592 & -1.0841208 & -0.9533426 & 0.0801835 \\
8 & -1.0613713 & -0.8751219 & -1.0741555 & -0.9533426 & 0.0782208 \\
9 & -1.0622522 & -0.9739388 & -1.0481834 & -0.9533426 & -0.0205962 \\
10 & -0.9293998 & -0.8287910 & -0.9526105 & -0.9533426 & 0.1245517 \\

\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.9}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-diff-p-tk-1.pdf}}

}

\caption{\label{fig-diff-p-tk}Posterior distribution of the difference
between the expected voicing proportion of /t, k/ vs /p/, averaged
across vowels.}

\end{figure}%

Once we have the posterior difference, we can obtain CrIs of the
difference using \texttt{quantile2()} from the posterior package. Beware
that the values of the difference are in log-odds! We can convert these
into odd-ratios with \texttt{exp()}. The output of the code below is
shown in Table~\ref{tbl-voi-bm-quant}. Odd-ratios indicate the ratio of
the difference between A and B, so that 1 means no difference, values
greater than 1 indicate an increase in A relative to B and values lower
than 1 indicate a decrease in A relative to B. odd-ratios are useful
when looking at differences that are in log-odds because while the
relative magnitude of the difference in proportion between two groups is
the same independent of the baseline proportion, the \emph{absolute}
magnitude of the difference depends on the baseline value. For example,
an odd-ratio difference of 1.25 would correspond to a proportion
increase of 13 percentage points if the baseline proportion is 0.62 but
it would correspond to a proportion increase of 17 percentage points if
the baseline proportion is 0.73. Of course, in real research contexts it
is still usefull to think about absolute magnitudes and their relevance
from a conceptual and methodological perspective. In this tutorial we
just focus on odd-ratios for simplicity.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(posterior)}

\NormalTok{voi\_bm\_quant }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{c2p\_tk\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# round to 2 digits}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrr@{}}

\caption{\label{tbl-voi-bm-quant}Upper and lower limits of 90, 80 and
60\% Credible Intervals of the difference ratio of voicing proportion in
/t, k/ vs /p/.}

\tabularnewline

\toprule\noalign{}
q90 & q80 & q60 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1.07 & 1.10 & 1.14 \\
1.38 & 1.34 & 1.30 \\

\end{longtable}

Based on the model and data, there is a 90\% probability that the
voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38\% increase)
than in /t, k/. At 80\% confidence, the change ratio is 1.10-1.34 (or
10-34\% increase) while at 60\% confidence is 1.14-1.30 (14-30\%
increase). In other words we can be quite confident that the voicing
proportion in /p/ is longer than in /t, k/ and that the increase is less
than 35\%. The brms package comes with a convenient function,
\texttt{conditional\_effects()}, to plot posterior means and CrIs based
on predictors in the model. In Figure~\ref{fig-voi-prop-bm-cond}, we
plot the predicted proportion of voicing by consonant and vowel.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{conditional\_effects}\NormalTok{(voi\_prop\_bm, }\StringTok{"c2:vowel"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-cond-1.pdf}}

}

\caption{\label{fig-voi-prop-bm-cond}Expected voicing proportion by
consonant and vowel with 95\% Credible Intervals.}

\end{figure}%

Finally, the package marginaleffects {[}XXX{]} has two other convenience
functions that return CrIs of comparisons across predictor levels
(\texttt{avg\_comparisons()}) and CrIs of posterior predictions across
predictor levels (\texttt{avg\_predictions}). {[}XXX{]}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(marginaleffects)}

\FunctionTok{avg\_comparisons}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \FunctionTok{list}\NormalTok{(}\AttributeTok{c2 =} \StringTok{"pairwise"}\NormalTok{), }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{type =} \StringTok{"link"}\NormalTok{)}
\FunctionTok{avg\_predictions}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \StringTok{"vowel"}\NormalTok{, }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}llrrr@{}}

\caption{\label{tbl-marginal-1}Difference in expected voicing proportion
by consonant, averaged across vowels, with 80\% CrIs.}

\tabularnewline

\toprule\noalign{}
term & contrast & estimate & conf.low & conf.high \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
c2 & mean(t) - mean(k) & 0.035 & -0.004 & 0.074 \\
c2 & mean(p) - mean(k) & 0.218 & 0.178 & 0.256 \\
c2 & mean(p) - mean(t) & 0.183 & 0.145 & 0.219 \\

\end{longtable}

\begin{longtable}[]{@{}lrrr@{}}

\caption{\label{tbl-marginal-2}Expected voicing proportion by vowel,
averaged across vowels, with 80\% CrIs.}

\tabularnewline

\toprule\noalign{}
vowel & estimate & conf.low & conf.high \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
i & 0.331 & 0.325 & 0.338 \\
e & 0.300 & 0.293 & 0.307 \\
a & 0.245 & 0.238 & 0.252 \\
o & 0.305 & 0.298 & 0.312 \\
u & 0.348 & 0.341 & 0.354 \\

\end{longtable}

\section{Case study 2: coarticulatory vowel
nasalisation}\label{sec-case-2}

For the second case study we will use data from \citet{carignan2021}.
The study looked at properties of nasality in German VNC sequences.
Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced
/d/) on the proportion of nasalisation within the vowel in the VNC
sequence. Previous work on coarticulatory nasalisation in English has
suggested that vowels followed by an NC sequence where C is voiceless
(NT) should show earlier coarticulatory nasalisation than vowels
followed by an NC sequence where C is voiced \citep[ND, see review
in][]{carignan2021}. This pattern has been suggested to be driven by the
articulatory and acoustic incompatibility of voicelessness and
nasalisation, by which the velum opening gesture of the nasal consonant
is pushed away (i.e.~earlier) when the consonant following the nasal is
voiceless. Everything else being equal, a greater proportion of vowel
nasalisation (from the perspective of time) should be found in vowels
followed by NT than in vowels followed by ND.

We will model the proportion of coarticulatory nasalisation in the
German short vowels /i, e, a, o, u/ when followed by /nt/ or /nd/, using
a Bayesian beta regression model. The proportion was calculated as the
proportion of the nasal interval to the duration of the vowel. The nasal
interval was defined thus: the interval between the time of peak
velocity of velum opening to the offset of the vowel. We will use the
results of the regression model to answer the following questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Is the nasalisation proportion, on average across vowels, greater in
  voiceless NC sequences?
\item
  Is there individual speaker variation?
\end{enumerate}

Table~\ref{tbl-nasal} shows the first ten rows of the data. The data
contains the following columns:

\begin{itemize}
\item
  \texttt{speaker} indicates the speaker ID.
\item
  \texttt{label} is the word label as given in the original data.
\item
  \texttt{vowel} is the target vowel in the VNC sequence.
\item
  \texttt{NC} is the NC sequence.
\item
  \texttt{voicing} indicates the voicing of C.
\item
  \texttt{nas\_prop} is the proportion of coarticulatory nasalisation of
  the vowel.
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lllllr@{}}

\caption{\label{tbl-nasal}First ten rows of the nasal proportion data.}

\tabularnewline

\toprule\noalign{}
speaker & label & vowel & NC & voicing & nas\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
S03 & b\_\emph{U\_nt}@\_\_\_N\_B17/s & u & nt & voiceless & 0.3668820 \\
S03 & b\_\emph{a\_nd}@\_\_\_N\_B19/s & a & nd & voiced & 0.1954858 \\
S03 & b\_\emph{a\_nt}@\_\_\_N\_B15/s & a & nt & voiceless & 0.2786485 \\
S03 & f\_\emph{I\_nt}@\_\_\_N\_B05/s & i & nt & voiceless & 0.7642259 \\
S03 & l\_\emph{I\_nd}@\_\_\_N\_B06/s & i & nd & voiced & 0.0052949 \\
S03 & p\_\emph{E\_nt}\_\_\_\_N\_B09/s & e & nt & voiceless &
0.3347331 \\
S03 & r\_\emph{a\_nt}@\_\_\_N\_B06/s & a & nt & voiceless & 0.2431760 \\
S03 & v\_\emph{I\_nd}@\_\_\_N\_B07/s & i & nd & voiced & 0.0247572 \\
S03 & v\_\_I\_nt\_6\_\_\_N\_B15/s & i & nt & voiceless & 0.1350081 \\
S03 & z\_\emph{E\_nd}@\_\_\_N\_B17/s & e & nd & voiced & 0.5378522 \\

\end{longtable}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nasal-1.pdf}}

}

\caption{\label{fig-nasal}Proportion of coarticulatory nasalisation
during the vowel in VNC sequences in German, depending on C voicing.}

\end{figure}%

Figure~\ref{fig-nasal} shows the proportion of coarticulatory
nasalisation in vowels followed by /nd/ (voiced) vs /nt/ (voiceless)
sequences, for the short vowels /i, e, a, o, u/. We can see a pattern of
higher nasalisation proportion in vowels followed by /nt/, at least in
the vowels /a, i, o/. For /e, u/, the distribution of nasalisation
proportion seems to be similar between the voiced and voiceless
contexts.

Now onto modelling with a beta regression. Note that a full appropriate
model would include further predictors (both constant and varying), but
for simplicity here we include only the following predictors: voicing
(voiced /nd/ vs voiceless /nt/), vowel (/i, e, a, o, u/), including an
interaction between them. As varying terms, we include a varying
intercept by speaker and a by-speaker varying slope for voicing and
vowel in interaction. As with the model from the first case study,
voicing and vowel are coded using indexing, by suppressing the intercept
with \texttt{0\ +}. Here's the code of the model.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{  nas\_prop }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{|}\NormalTok{ speaker),}
  \AttributeTok{data =}\NormalTok{ nasal,}
  \AttributeTok{family =}\NormalTok{ Beta,}
  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
  \AttributeTok{file =} \StringTok{"data/cache/nas\_prop\_bm"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Let's inspect the output of \texttt{fixef()}, reported in
Table~\ref{tbl-nas-prop-fixef}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-nas-prop-fixef}Regression coefficients of a beta
regression of nasalisation proportion (\texttt{nas\_prop\_bm}).}

\tabularnewline

\toprule\noalign{}
& Estimate & Est.Error & Q10 & Q90 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
voicingvoiced:vowela & -0.61 & 0.12 & -0.77 & -0.46 \\
voicingvoiceless:vowela & -0.08 & 0.09 & -0.19 & 0.04 \\
voicingvoiced:vowele & -0.19 & 0.16 & -0.38 & 0.01 \\
voicingvoiceless:vowele & -0.25 & 0.10 & -0.38 & -0.12 \\
voicingvoiced:voweli & -0.74 & 0.19 & -0.98 & -0.51 \\
voicingvoiceless:voweli & -0.12 & 0.18 & -0.35 & 0.10 \\
voicingvoiced:vowelo & -0.34 & 0.16 & -0.54 & -0.14 \\
voicingvoiceless:vowelo & 0.25 & 0.15 & 0.05 & 0.44 \\
voicingvoiced:vowelu & -0.71 & 0.18 & -0.94 & -0.49 \\
voicingvoiceless:vowelu & -0.58 & 0.16 & -0.78 & -0.38 \\

\end{longtable}

Negative log-odds indicate a proportion that is smaller than 50\%, while
positive log-odds a proportion that is greater than 50\%. Generally, the
expected log-odd predictions are negative, indicating an overall
tendency for the nasalisation to take less than 50\% of the duration of
the vowel. Moreover, the predictions are higher for voiceless NC
sequences than for voiced NC sequences, indicating a greater proportion
of nasalisation in the former. However there is vowel-specific
variation, and there doesn't seem to be much of a difference in
nasalisation proportion in /e/ and /u/. Figure~\ref{fig-nas-prop} shows
the expected predictions with \texttt{conditional\_effects()}, which
should make the interpretation clearer.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{conditional\_effects}\NormalTok{(nas\_prop\_bm, }\StringTok{"vowel:voicing"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-1.pdf}}

}

\caption{\label{fig-nas-prop}Expected nasalisation proportion by vowel
and voicing, with 95\% CrIs.}

\end{figure}%

Now that we fitted the model we can use the draws to answer the two
research questions (repeated from above):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Is the nasalisation proportion, on average across vowels, greater in
  voiceless NC sequences?
\item
  Is there individual speaker variation?
\end{enumerate}

To answer question 1, we can calculate the average difference in
nasalisation proportion by first calculating the average nasalisation
across all vowels for voiced and voiceless sequences (see
Table~\ref{tbl-nas-draws} for the output of this step) and then take the
difference of those, similarly to what we have done in the Case Study 1
above.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{nas\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(nas\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}

\NormalTok{nas\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot vowel:c2 columns}
  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiced:vowela}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiceless:vowelu}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrlllr@{}}

\caption{\label{tbl-nas-draws}First ten rows of posterior draws from
\texttt{nas\_prop\_bm} in long format.}

\tabularnewline

\toprule\noalign{}
.chain & .iteration & .draw & type & voicing & vowel & value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 1 & 1 & b & voicingvoiced & vowela & -0.4632870 \\
1 & 1 & 1 & b & voicingvoiceless & vowela & -0.0943699 \\
1 & 1 & 1 & b & voicingvoiced & vowele & -0.1546258 \\
1 & 1 & 1 & b & voicingvoiceless & vowele & -0.2080810 \\
1 & 1 & 1 & b & voicingvoiced & voweli & -0.4810383 \\
1 & 1 & 1 & b & voicingvoiceless & voweli & -0.0529377 \\
1 & 1 & 1 & b & voicingvoiced & vowelo & -0.3096219 \\
1 & 1 & 1 & b & voicingvoiceless & vowelo & 0.3212456 \\
1 & 1 & 1 & b & voicingvoiced & vowelu & -0.6198312 \\
1 & 1 & 1 & b & voicingvoiceless & vowelu & -0.3680900 \\

\end{longtable}

Now let's calculate the mean nasalisation proportion within each draw by
voicing, and plot the resulting posterior distributions. Note that, as
discussed for Case study 1, when working wit log-odds it is important to
first do all necessary calculations in log-odds, here calculate the mean
log-odds across vowels, and \emph{then} transform the calculated
estimands to proportions/probabilities. The output of the following code
is show in Table~\ref{tbl-nas-long-voi} and the density plot of the
calculated draws is in
Figure~\ref{fig-nas-prop-bm-draws-long-voicing-plot}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and voicing ensures that averaging applies only within draw and voicing}
  \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{summarise}\NormalTok{(}
    \CommentTok{\# calculate the mean value within draw/voicing in log{-}odds}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value),}
    \CommentTok{\# we can now transform log{-}odds to proportion with plogis()}
    \AttributeTok{value\_mean\_prop =} \FunctionTok{plogis}\NormalTok{(value\_mean),}
    \AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rlrr@{}}

\caption{\label{tbl-nas-long-voi}Mean nasalisation proportion by draw
and voicing, averaged across vowels (first 10 rows).}

\tabularnewline

\toprule\noalign{}
.draw & voicing & value\_mean & value\_mean\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & voicingvoiced & -0.4056808 & 0.3999482 \\
1 & voicingvoiceless & -0.0804466 & 0.4798992 \\
2 & voicingvoiced & -0.3905744 & 0.4035790 \\
2 & voicingvoiceless & -0.1227124 & 0.4693603 \\
3 & voicingvoiced & -0.4436082 & 0.3908816 \\
3 & voicingvoiceless & -0.1320190 & 0.4670431 \\
4 & voicingvoiced & -0.4385607 & 0.3920840 \\
4 & voicingvoiceless & -0.0981768 & 0.4754755 \\
5 & voicingvoiced & -0.3827134 & 0.4054726 \\
5 & voicingvoiceless & -0.2657250 & 0.4339569 \\

\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(value\_mean\_prop, voicing)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-bm-draws-long-voicing-plot-1.pdf}}

}

\caption{\label{fig-nas-prop-bm-draws-long-voicing-plot}Expected
nasalisation proportion by voicing, averaged across vowels.}

\end{figure}%

The plot suggests an overall greater nasalisation proportion in
voiceless NC sequences. Let's quantify how greater as we did in Case
Study 1. We will use odd-ratios in this context as well, i.e.~we will
convert log-odds to odd-ratios using the \texttt{exp()} (exponential)
function (and as before we first calculate the difference and then
exponentiate the resulting values, after which we can take summary
measures, like means and quantile-based measures such as CrIs).
tbl-nas-prop-bm-diff-quant show the 90, 80 and 60\% CrIs of the
difference ratio of nasalisation proportion in voiceless vs voiced NC
sequences.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
  \CommentTok{\# with one draw per raw. we need to drop the value\_mean\_prop col}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{value\_mean\_prop) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the difference of voiceless and voiced in log{-}odds}
    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced,}
    \CommentTok{\# now transform with exp() to get the ratio difference}
    \AttributeTok{voicing\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(voicing\_diff)}
\NormalTok{  )}

\NormalTok{nas\_prop\_bm\_diff\_quant }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrr@{}}

\caption{\label{tbl-nas-prop-bm-diff-quant}Upper and lower limits of 90,
80 and 60\% Credible Intervals of the difference ratio of nasalisation
proportion in vowels followed by voiced vs voiceless consonants.}

\tabularnewline

\toprule\noalign{}
q90 & q80 & q60 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1.23 & 1.27 & 1.33 \\
1.69 & 1.63 & 1.56 \\

\end{longtable}

The CrIs of the ratio difference in nasalisation proportion in voiceless
vs voiced NC sequences suggest an increase of nasalisation in the
voiceless NC sequences, with a 90\% probability that the increase is
between 23\% and 69\% of the proportion in voiced NC sequences.

Moving onto question 2: is there individual speaker variation?
{[}TAMMINGA XXX{]} For this, we will use the \texttt{spread\_draws()}
function from tidybayes {[}XXX{]} to extract the draws of the varying
terms (in brms these are the coefficients that start with \texttt{r\_}).
There is quite a few steps of processing to get from the raw draws to
the estimand we need: while we have commented the following code, we
encourage readers to test each line sequentialy and inspect the
intermediate output to fully understand the process. We assume that
readers are familiar enough with models with varying terms (aka random
effects, mixed-effects models). What readers should note is that to
obtain the expected predictions of nasalisation proportion for each
speaker, the constant terms and the varying terms should be added (since
the varying terms indicate the deviation of each speaker from the
overall estimate).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidybayes)}

\NormalTok{nas\_prop\_r }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# extract draws from model, only \textasciigrave{}r\_speaker\textasciigrave{} varying terms}
  \FunctionTok{spread\_draws}\NormalTok{(r\_speaker[speaker,voicingvowel]) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# separate the column voicingvowel to two columns}
  \FunctionTok{separate}\NormalTok{(voicingvowel, }\FunctionTok{c}\NormalTok{(}\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# join the draws with the \textasciigrave{}b\_\textasciigrave{} terms}
  \FunctionTok{left\_join}\NormalTok{(}\AttributeTok{y =}\NormalTok{ nas\_prop\_bm\_draws\_long) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# get the expected log{-}odd value of each speaker, in each draw}
  \CommentTok{\# this is the sum of the \textasciigrave{}value\textasciigrave{} from the b\_ terms and the value from the}
  \CommentTok{\# r\_speaker term.}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{r\_speaker\_value =}\NormalTok{ value }\SpecialCharTok{+}\NormalTok{ r\_speaker) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# group the data for summarise}
  \FunctionTok{group\_by}\NormalTok{(.draw, speaker, voicing) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# get mean expected log{-}odds by draw, speaker and voicing (averaging across vowel)}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{r\_speaker\_value\_mean =} \FunctionTok{mean}\NormalTok{(r\_speaker\_value)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# make the data wider: two columns, one for voiced and one for voiceless}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ r\_speaker\_value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# finally, calculate the difference in expected log{-}odds of voiceless and voiced}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced)}
\end{Highlighting}
\end{Shaded}

Figure~\ref{fig-nas-prop-r} plots the posterior distributions of the
expected log-odd difference of coarticulatory nasalisation in voiceless
vs voiced NC sequences (\emph{x}-axis), for each speaker in the data
(\emph{y}-axis), as predicted by the model. The red solid vertical line
indicates the constant (overall) expected log-odd difference based on
the (constant) \texttt{b\_} terms. The black dashed vertical line marks
log-difference 0 (i.e., no difference in proportion of nasalisation
between voiceless and voiced NC).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_r }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(voicing\_diff, }\FunctionTok{reorder}\NormalTok{(speaker, voicing\_diff))) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \FunctionTok{mean}\NormalTok{(nas\_prop\_bm\_diff}\SpecialCharTok{$}\NormalTok{voicing\_diff), }\AttributeTok{colour =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{lims}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\FloatTok{1.5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-r-1.pdf}}

}

\caption{\label{fig-nas-prop-r}Expected log-odd difference of
coarticulatory nasalisation in voiceless vs voiced NC sequences for each
speaker.}

\end{figure}%

There is a lot of uncertainty within and between speakers: while the
distributions of most speakers are located in the positive range, some
expected distributions (see last 5 speakers at the bottom of figure) do
substantially span both negative and positive values. In other words,
while most speakers are more likely to have a larger nasalisation
proportion in voiceless NC sequences, a few might in fact have the
opposite pattern. Even among those speakers that do have a more robust
positive difference, there is a lot of uncertainty as to the magnitude
of the difference.

  \bibliography{linguistics.bib,statistics.bib}

\end{document}