### 2025-02-10 manuscript/manuscript.tex
add example of bad gaussian

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}

\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A tutorial of Bayesian beta regressions with brms in R},
  pdfauthor={Stefano Coretta; Paul Bürkner},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{A tutorial of Bayesian beta regressions with brms in R}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk} 
\author[]{Paul Bürkner%
%
}
 \ead{paul.buerkner@gmail.com}

\affiliation[1]{organization={University of Edinburgh, Linguistics and
English Language},addressline={3 George
Sq},city={Edinburgh},postcode={EH8 9AD},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{introduction}

Phonetic research often involves numeric continuous outcome variables,
like durations, frequencies, loudness and ratios
\citep{kluender1988, johnson2003, gick2013, ladefoged2014, abramson2017, roettger2018, coretta2023, rosen2005}.
Another commonly employed type of outcome variable are proportions: for
example, proportion of voicing during closure \citep{davidson2016},
vocal folds contact quotient \citep{herbst2017}, gesture amplitude
\citep{carignan2021}, nasalance \citep{carignan2021}. Moreover,
virtually any measure can be MIN-MAX normalised, a procedure which
transforms values so that they are in the range 0--1.

Regression models (and extensions like generalised additive models) have
become a \emph{de facto} standard for the statistical analysis of a
variety of measures in phonetic research
\citep{kirby2018, politzerahles2018, tavakoli2024}. However, there is a
tendency for researchers to use Gaussian distribution families
(i.e.~probability distributions for the outcome variable) for any
measure that is numeric, irrespective of whether the measure is
unbounded, like in truly Gaussian variables, or bounded, like in
proportions. A possible reason is that the base R function for fitting
regression models, \texttt{lm()}, and the lme4 function used to fit
regression models with varying terms, \texttt{lmer()} from lme4
\citep{bates2015}, both fit Gaussian regressions by default and the user
does not have to specify the distribution family. This tacit default of
using Gaussian models is also reflected in teaching practices, where
significance test and models using the Gaussian distribution are the
first to be taught \citep{baayen2008, winter2020}, due to their relative
simplicity and the fact that regression models with other families are
conceptual generalisations of Gaussian regression models.
Figure~\ref{fig-bg} shows how poorly a Gaussian regression model fits
beta-distributed data: the dark blue line is the density of data
simulated to be distributed according to a beta distribution; the light
blue lines are predicted distributions as generated by a Gaussian model
of the data. The predicted Gaussian distributions over-predict values
between 0.5 and 0.75 and under-predict values between 0.75 and 1.
Moreover, the density peak (median) of the Gaussian distributions is
lower than the peak in the original data.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-bg-1.pdf}}

}

\caption{\label{fig-bg}Gaussian model fitted to beta-distributed data.
The dark blue line is the density of the beta data, while the light blue
lines are simulated densities of the same data based on a Gaussian
model.}

\end{figure}%

While most researchers approach proportions with Gaussian regression
models, proportion are not Gaussian by nature, since they are continuous
variables bounded between 0 and 1. Thus, regression models with
proportions as outcome variable should be fitted using a likelihood
family that assumes such unit interval data. A common and flexible
distribution for this purpose is the beta distribution. This tutorial
introduces researchers to beta regression models in R using the package
brms. Familiarity with regression modelling in R with a package like
lme4 is assumed, but no prior knowledge of Bayesian statistics is
necessary. The rest of the paper is structured as follows:
Section~\ref{sec-beta} introduces the mathematical underpinnings of the
beta distribution, Section~\ref{sec-bayes} provides the readers with a
brief explanation and justification of Bayesian regression models, while
Section~\ref{sec-case-1} and Section~\ref{sec-case-2} illustrate how to
fit Bayesian beta regression models with two case studies.

\section{The beta distribution}\label{sec-beta}

The beta distribution is often considered as \emph{the} distribution for
the modelling of unit interval data
\citep{ferrari_beta_2004, cribari-neto_beta_2010}. It has been
extensively studied theoretically
\citep{krysicki_new_1999, gupta_handbook_2014, espinheira_beta_2008} and
is used a baseline to compare other unit interval distributions against
\citep{kieschnick_regression_2003, bonat_regression_2013, lopez_bayesian_2013}.
Specifically in Bayesian regression models, it was shown to perform well
across a wide range of scenarios
\citep{scholz_posterior_2023, scholz_prediction_2025}, which is why we
focus on the beta distribution in this tutorial.

We use a common mean parametrisation for the beta distribution with mean
parameter \(\mu\), bounded between 0 and 1, and positive precision
parameter \(\phi\) that is roughly proportional to the inverse of the
variance \(\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}\). That is,
the larger \(\phi\) the smaller the variance of the corresponding beta
distribution. While the mathematical details are not needed to
understand the content of this tutorial, we still write down the density
below for reasons of completeness:

\[
p(y \mid \mu, \phi) = \frac{1}{B(\mu \, \phi, \; (1-\mu) \, \phi)} \; y^{\mu \, \phi -1} \; (1-y)^{(1-\mu) \, \phi -1},
\]

where \(B\) is the Beta function, a complex integral for whose numerical
approximation efficient algorithms exist in every programming language.

\section{Bayesian regression models}\label{sec-bayes}

Bayesian regression models are being increasingly adopted within
phonetics and language research more broadly
\citep{vasishth2018, nalborczyk2019, verissimo2021}. Bayesian inference
involves updating of prior probability distributions in light of
evidence from data, to produce posterior probability distributions. In
Bayesian regression models, model parameters are modelled as full
probability distributions, rather than point estimates as in Null
Hypothesis Significance Testing. Given the difficulty of analytical
solutions of model equations, Bayesian regressions rely on sampling
algorithms to reconstruct the posterior distributions. The statistical
language Stan \citep{standevelopmentteam2017} employs efficient and
robust Markov Chain Monte Carlo algorithms for fitting a variety of
models, and the R package brms allows R users to interface with Stan
from within R to fit Bayesian regression models
\citep{burkner2017, burkner2018, burkner2021}.

The main practical advantage of Bayesian regression models over
maximum-likelihood-based frequentist regression models, like those
fitted with the lme4 package \citep{bates2015}, is that Bayesian
regression models don't suffer from the convergence issues that models
fitted in lme4 \citep{bates2015} and other packages
\citep{cribarineto2010} that fit frequentist models do, independent of
sample size. A second, long-term advantage is that Bayesian regression
models allow researchers to statistically re-use information from
previous studies by specifying informative priors. While prior
specification is one of the main features of Bayesian inference, in this
tutorial we will use the default priors as set by brms. These are
sensible priors estimated from the data that facilitate convergence but
bear virtually no influence on the estimated posteriors. Specifying
priors requires a great deal of precise quantitative knowledge, which in
most areas of phonetics we still do not possess, so that using default
uninformative priors is, for the time being, theoretically sound. For
practical guidance on prior specification, see
\url{https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations}.

The output of Bayesian regression models is (posterior) probability
distributions for the model parameters, through which researchers can
quantify (un)certainty. Bayesian Credible Intervals (CrIs) can be
calculated from the posterior distributions at several probability
levels (e.g.~95, 90, 80, 60\%) for a more complete view on estimated
parameters. Bayesian CrIs indicate that at a certain probability levels
the parameter's estimate lies within that interval: so, for example, a
90\% CrI {[}A, B{]} indicates that there is a 90\% probability that the
estimate is between A and B. Different probability levels correspond to
different levels of confidence: the higher the probability the higher
the confidence (always conditional on data and model). Readers
interested in a full and accessible exposition of Bayesian statistics
are referred to \citet{mcelreath2019}. Shorter introductions can be
found in \citet{etz2018}, \citet{vasishth2018} and
\citet{nalborczyk2019}.

\section{Case study 1: coarticulatory voicing within consonant
closure}\label{sec-case-1}

For the first case study, we will model the proportion of coarticulatory
progressive voicing within the consonant closure of voiceless stops. The
measurements come from a data set of audio and electroglottographic
(EGG) recordings of 19 speakers of Northwestern Italian
\citep{coretta2019k, coretta2020b}. The participants read frame
sentences which included target words of the form /CVCo/, where C was
either /k, t, p/ in all permutations and V was either /i, e, a, o, u/
(two resulting words, /peto/ and /kako/ were excluded because they are
profanities), for a total of 43 target words. There were 4 different
frame sentence: \emph{Scrivete X sul foglio} `Write X on the sheet',
\emph{Ha detto X sei volte} `She said X six times', \emph{Sentivo X di
nuovo} `I heard X again', \emph{Ripete X da sempre} `She's been
repeating X since ever'. There is a total of 172 trials per participant
(3,268 grand total). The actual observation count is 2,419, after
removing speech errors, EGG measurement errors, and cases in which
voicing ceased before the closure onset/after the closure offset of the
second consonant in the target word.

The proportion of voicing during the closure of the second consonant was
calculated as the proportion of contiguous voicing duration after
closure onset to total duration of closure. The following code chunk
attaches the tidyverse packages \citep[for reading and wrangling
data,][]{wickham2019} and loads the \texttt{ita\_egg} tibble (data
frame). The tibble is filtered so as to remove voicing proportions
(\texttt{voi\_clo\_prop}) that are smaller than 0 and greater than 1.
The variables \texttt{vowel} (first vowel in the target word) and
\texttt{c2} (second consonant in the target word) are converted to
factors to specify the order of the levels. Table~\ref{tbl-ita-egg}
shows the first ten rows of the tibble (only relevant columns are
included).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach tidyverse and set light theme for plots}
\FunctionTok{library}\NormalTok{(tidyverse)}
\FunctionTok{theme\_set}\NormalTok{(}\FunctionTok{theme\_light}\NormalTok{())}

\CommentTok{\# load tibble}
\FunctionTok{load}\NormalTok{(}\StringTok{"data/coretta2018/ita\_egg.rda"}\NormalTok{)}

\CommentTok{\# filter and mutate data}
\NormalTok{ita\_egg }\OtherTok{\textless{}{-}}\NormalTok{ ita\_egg }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{filter}\NormalTok{(voi\_clo\_prop }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{, voi\_clo\_prop }\SpecialCharTok{\textless{}} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{vowel =} \FunctionTok{factor}\NormalTok{(vowel, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"i"}\NormalTok{, }\StringTok{"e"}\NormalTok{, }\StringTok{"a"}\NormalTok{, }\StringTok{"o"}\NormalTok{, }\StringTok{"u"}\NormalTok{)),}
    \AttributeTok{c2 =} \FunctionTok{factor}\NormalTok{(c2, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"k"}\NormalTok{, }\StringTok{"t"}\NormalTok{, }\StringTok{"p"}\NormalTok{))}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}llllr@{}}

\caption{\label{tbl-ita-egg}First 10 rows of the data of voicing
proportion within stop closure.}

\tabularnewline

\toprule\noalign{}
speaker & word & vowel & c2 & voi\_clo\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
it01 & poto & o & t & 0.0457950 \\
it01 & topo & o & p & 0.3522513 \\
it01 & pato & a & t & 0.1440749 \\
it01 & teto & e & t & 0.3610916 \\
it01 & toto & o & t & 0.2743519 \\
it01 & puco & u & k & 0.1943984 \\
it01 & chipo & i & p & 0.2227896 \\
it01 & peco & e & k & 0.1835596 \\
it01 & poco & o & k & 0.1771007 \\
it01 & poto & o & t & 0.1525163 \\

\end{longtable}

Figure~\ref{fig-ita-egg} shows the raw voicing duration proportion
values split by vowel /i, e, a, o, u/ and second consonant /k, t, p/ in
the /CVCo/ target words. The plot suggests that, on average, the voicing
proportion is slightly lower with /k/ than with /t, p/. Moreover, there
is greater variability between vowels in /t, p/ than in /k/. We will use
a beta regression model to assess whether the place of articulation of
the consonant affects the proportion of coarticulatory progressive
voicing, while averaging across vowels (of course, another question that
could be answered with the same model is if different vowels lead to
difference voicing proportions, but for sake of space we restrict this
tutorial to the former question). A general expectation is that the
proportion of progressive voicing should increase with fronter places of
articulation (so from /k/ to /t/ to /p/) given the aerodynamic
constraints which obtain with vocal fold vibration during a fully
occluded oral tract \citep{vandenberg1958, rothenberg1967, ohala2011}.

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-ita-egg-1.pdf}}

}

\caption{\label{fig-ita-egg}Proportion of voicing during the closure of
the second consonant in /CVCo/ words by vowel and the second consonant.}

\end{figure}%

We will use brms to fit a Bayesian beta regression \citep{burkner2017}.
The model has voicing proportion as the outcome variable and the
following terms: an interaction between vowel (/i, e, a, o, u/) and
second consonant C2 (/k, t, p/), centred speech rate (number of
syllables per second); as varying (aka random) terms, we included
by-speaker varying coefficients for the vowel/consonant interaction and
for centred speech rate.\footnote{While the terms ``random effects,
  intercept and slopes'' are commonly employed in applied statistical
  work in linguistics and phonetics, we opt to use the conceptually more
  appropriate terminology proposed by \citet{gelman2005}: varying terms,
  intercepts, slopes, coefficients. Other common terms are multilevel
  and hierarchical terms.} The categorical predictors vowel and C2 are
coded using indexing rather than the traditional R contrasts, as
advocated in \citet{mcelreath2019}: in R, this corresponds to
suppressing the model's intercept with the \texttt{0\ +} syntax; using
indexing instead of contrasts makes it easier to specify priors and
uncertainty is distributed equally across all levels of categorical
predictors. For pedagogical simplicity, the model will use the default
priors as explained in Section~\ref{sec-bayes}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach brms}
\FunctionTok{library}\NormalTok{(brms)}

\CommentTok{\# fit the model}
\CommentTok{\# Takes 3 minutes on MacBook Pro 2020, M1}
\NormalTok{voi\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
  \CommentTok{\# model formula}
\NormalTok{  voi\_clo\_prop }\SpecialCharTok{\textasciitilde{}}
    \CommentTok{\# constant terms}
    \DecValTok{0} \SpecialCharTok{+}\NormalTok{ vowel}\SpecialCharTok{:}\NormalTok{c2 }\SpecialCharTok{+}\NormalTok{ speech\_rate\_c }\SpecialCharTok{+}
    \CommentTok{\# varying terms}
\NormalTok{    (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ vowel}\SpecialCharTok{:}\NormalTok{c2 }\SpecialCharTok{+}\NormalTok{ speech\_rate\_c }\SpecialCharTok{|}\NormalTok{ speaker),}
  \CommentTok{\# uses the beta family for the outcome}
  \AttributeTok{family =}\NormalTok{ Beta,}
  \AttributeTok{data =}\NormalTok{ ita\_egg,}
  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
  \AttributeTok{file =} \StringTok{"data/cache/voi\_prop\_bm"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. The full summary with an explanation of each
part can be found in XXX. Table~\ref{tbl-voi-prop-fixef} reports the
output of \texttt{fixef()} as a table (we round all values to the
nearest 2 digits for clarity.). For each coefficient in the model,
\texttt{fixef()} prints the name of the coefficient, the mean estimate,
the estimate error and the lower and upper limits of a Bayesian Credible
interval (CrI). Note that all values are in log-odds, since the link
function in beta regressions is the logistic function, like in
binomial/Bernoulli regressions. Here, we print an 80\% CrI. There is
nothing special about 95\% CrI within Bayesian inference and instead
experts recommend to check and report a variety of CrIs. Obtaining CrIs
at different probability levels allows researchers to make more
fine-grained inferential statements than the frequentist significance
dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
in this case study but we strongly recommend researchers to always
obtain CrIs at different levels of probability and base their inferences
on all and not one in particular. To reiterate, in Bayesian inference,
an 80\% CrI indicates the range of values within which the true estimate
falls at 80\% probability or confidence.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-voi-prop-fixef}Regression coefficients of a beta
regression of voicing proportion (\texttt{voi\_prop\_bm}).}

\tabularnewline

\toprule\noalign{}
& Estimate & Est.Error & Q10 & Q90 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
speech\_rate\_c & 0.08 & 0.06 & 0.01 & 0.15 \\
voweli:c2k & -0.91 & 0.14 & -1.08 & -0.74 \\
vowele:c2k & -1.08 & 0.11 & -1.22 & -0.94 \\
vowela:c2k & -0.99 & 0.12 & -1.14 & -0.84 \\
vowelo:c2k & -0.79 & 0.14 & -0.96 & -0.62 \\
vowelu:c2k & -1.00 & 0.16 & -1.20 & -0.80 \\
voweli:c2t & -0.66 & 0.11 & -0.79 & -0.53 \\
vowele:c2t & -0.84 & 0.14 & -1.02 & -0.66 \\
vowela:c2t & -1.43 & 0.13 & -1.60 & -1.26 \\
vowelo:c2t & -1.15 & 0.13 & -1.31 & -0.99 \\
vowelu:c2t & -0.68 & 0.12 & -0.83 & -0.54 \\
voweli:c2p & -0.68 & 0.11 & -0.81 & -0.54 \\
vowele:c2p & -0.88 & 0.15 & -1.07 & -0.68 \\
vowela:c2p & -1.14 & 0.13 & -1.31 & -0.98 \\
vowelo:c2p & -0.66 & 0.11 & -0.80 & -0.53 \\
vowelu:c2p & -0.44 & 0.12 & -0.59 & -0.28 \\

\end{longtable}

From the summary, we see that speech rate (number of syllables per
second) has a positive effect on voicing proportion: the 80\% CrI is
between 0.01 and 0.15 log-odds {[}\(\beta\) = 0.08, SD = 0.06{]}.
Log-odds can be converted to odd-ratios by exponentiating the value:
0.01-0.15 log odds correspond to an odd-ratio of 1.01 to 1.16, or as
percentages, to an increase of voicing of 1 to 16\% for every increase
of one syllable per second. Since this is an 80\% CrI, we can be 80\%
confident that the true effect of speech rate is between 1-16\% increase
of voicing proportion, conditional on the data and model. Note that
transforming measures this way is appropriate \emph{only} with
quantile-based measures (like CrIs) but not with moments like the mean
and standard deviation: to correctly get mean and SDs in the transformed
scale, you must first extract the posterior draws (see below), convert
them and then take moments such as mean and SD \citep[for a more
detailed explanation, see][p.~101]{burkner2025}. In the avoidance of
doubt, we will always transform the drawn values first and then take
summary measures.

Turning now to the coefficients for vowel and C2, given the indexing
approach of coding these variables the model summary and the output of
\texttt{fixef()} reports the \emph{predictions} in log-odds for each
combination of vowel and C2, rather than differences between levels. The
CrIs of the vowel/C2 coefficients span all negative log-odds values:
these correspond to proportions that are lower than 0.5 (which is 0 in
log-odds). This matches the general trends in the raw data, which we
plotted in Figure~\ref{fig-ita-egg}. Next, we will plot the predicted
proportions of each vowel/C2 combination at mean speech rate
(i.e.~centred speech rate = 0) and then calculate the average pair-wise
difference in voicing proportion between /k, t, p/.

Before being able to plot the predictions, it's important to get
familiar with the so-called posterior draws. Bayesian regression models
are fitted using Markov Chain Monte Carlo (MCMC) methods, as briefly
explained in Section~\ref{sec-bayes}. The \texttt{brm()} function runs
three MCMC chains with 2000 iterations each by default. These are
sufficient in most cases. The first 1000 iterations of each chain are
used for ``warmup'', where the MCMC algorithm is optimised to find the
posterior probability. The second half of the iterations, 1000 per
chain, are used for sampling from the posterior probability. Each
iteration returns a drawn value for each parameter in the model. The
list of values drawn during the Monte Carlo sampling are called the
``posterior draws''. The posterior draws are used for inference: we can
plot them, summarise them, transform them and combine them in any way to
answer specific questions.

Posterior draws can be conveniently obtained with the
\texttt{as\_draws\_df()} from brms. For the moment, we will extract only
the draws of the constant regression coefficients (model variables
starting with \texttt{b\_}). To check which coefficients are available
in a model, use \texttt{get\_variables()} from the tidybayes package
\citep{kay2019}. \texttt{as\_draws\_df()} returns a tibble where each
column contains the drawn values of a coefficient.
Table~\ref{tbl-voi-prop-draws} shows the first ten rows and first five
columns of the output of \texttt{as\_draws\_df()}. The probability
distribution of the drawn values of each coefficient is the posterior
probability distribution of that coefficient. Note that, due to the
indexing coding of vowel and C2, all coefficient except
\texttt{b\_speech\_rate\_c} are \emph{predicted log-odds} for each
vowel/C2 combination (the drawn values for \texttt{b\_speech\_rate\_c}
are drawn \emph{differences} in log-odds for each unit increase of
speech rate). The drawn values are in log-odds, but we can convert them
to proportions with \texttt{plogis()} (we will do this when plotting
below).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{voi\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: Dropping 'draws_df' class as required metadata was removed.
\end{verbatim}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2353}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}@{}}

\caption{\label{tbl-voi-prop-draws}First ten rows and 5 columns of the
posterior draws for the model \texttt{voi\_prop\_bm}.}

\tabularnewline

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
b\_speech\_rate\_c
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_voweli:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowele:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowela:c2k
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
b\_vowelo:c2k
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.1593890 & -0.7213516 & -1.046577 & -1.1197185 & -1.0191525 \\
0.0351033 & -0.7492306 & -1.385450 & -1.0633164 & -0.8854320 \\
0.1259352 & -0.8472495 & -0.975776 & -1.0972735 & -0.6440836 \\
0.0640492 & -0.7033469 & -1.068626 & -1.0475190 & -0.8328252 \\
0.0319081 & -0.7659281 & -1.177656 & -0.9357499 & -0.6955130 \\
0.0729965 & -0.8494392 & -1.226692 & -1.0635364 & -0.7692243 \\
0.1009609 & -0.9546148 & -1.072955 & -1.1408255 & -0.8566750 \\
0.0735392 & -1.0160675 & -1.065025 & -1.1736263 & -0.8334884 \\
0.0870061 & -1.0436475 & -1.069337 & -1.2632061 & -0.7941857 \\
0.0923815 & -1.2042168 & -1.041238 & -0.9158689 & -0.6444492 \\

\end{longtable}

We can now wrangle this tibble and plot the posterior distributions for
each vowel/C2 combination. Table~\ref{tbl-voi-long} shows the first ten
rows.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# drop b\_speech\_rate\_c before pivoting}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{b\_speech\_rate\_c) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot vowel:c2 columns}
  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voweli:c2k}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_vowelu:c2p}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{, }\StringTok{"c2"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrlllr@{}}

\caption{\label{tbl-voi-long}First ten rows of posterior draws from
\texttt{voi\_prop\_bm} in long format.}

\tabularnewline

\toprule\noalign{}
.chain & .iteration & .draw & type & vowel & c2 & value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 1 & 1 & b & voweli & c2k & -0.7213516 \\
1 & 1 & 1 & b & vowele & c2k & -1.0465766 \\
1 & 1 & 1 & b & vowela & c2k & -1.1197185 \\
1 & 1 & 1 & b & vowelo & c2k & -1.0191525 \\
1 & 1 & 1 & b & vowelu & c2k & -1.0523007 \\
1 & 1 & 1 & b & voweli & c2t & -0.6248875 \\
1 & 1 & 1 & b & vowele & c2t & -0.7951552 \\
1 & 1 & 1 & b & vowela & c2t & -1.3999773 \\
1 & 1 & 1 & b & vowelo & c2t & -1.0950180 \\
1 & 1 & 1 & b & vowelu & c2t & -0.6682030 \\

\end{longtable}

For plotting, we can use ggplot2 statistics layers from the ggdist
package {[}\citet{kay2023}; kay2024{]}. \texttt{stat\_halfeye()} plots
the density of the posterior probability (in grey), its median (point)
and CrIs (lines). Let's use 60 and 80\% CrIs and transform the log-odds
values to proportions with \texttt{plogis()}. See
Figure~\ref{fig-voi-prop-bm-1}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# attach ggdist package}
\FunctionTok{library}\NormalTok{(ggdist)}

\NormalTok{voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value), vowel)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2)) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{caption =} \StringTok{"The points are the medians and the two lines are 60 and 80\% CrIs."}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-1-1.pdf}}

}

\caption{\label{fig-voi-prop-bm-1}Expected predictions of voicing
proportion by vowel and consonant from \texttt{voi\_prop\_bm}.}

\end{figure}%

What if we want to plot the average predicted voicing proportion for the
three consonants /k, t, p/? One approach is to take the mean across
vowels within each consonant for each posterior draw, and the posterior
distribution of the resulting list of values is the predicted posterior
distribution of voicing proportion for each consonant, assuming an
``average'' vowel.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and c2 ensures that averaging applies only within draw and c2}
  \FunctionTok{group\_by}\NormalTok{(.draw, c2) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# calculate the mean value within draw/c2}
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), c2)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{caption =} \StringTok{"The points are the medians and the two lines are 60 and 80\% CrIs."}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-pbdl-c2-1.pdf}}

}

\caption{\label{fig-voi-pbdl-c2}Expected predictions of voicing
proportion by consonant, averaged across vowels.}

\end{figure}%

Based on the expected posterior distributions of the mean voicing
proportion by consonant, /p/ has a somewhat higher voicing proportion
than /k/ and /t/. The real question is: how much higher? We can quantify
this by taking the difference of the drawn values for /p/ and those for
/t, k/ (all averaged across vowels). Since we want to compare /t, k/
with /p/, we should first average the average draws of /t, k/ and then
take the difference of the averaged /t, k/ draws and the draws of /p/.
Table~\ref{tbl-voi-pbd} shows the first ten rows of the resulting data
frame. The posterior distribution of the expected difference is shown in
Figure~\ref{fig-diff-p-tk}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per consonant with the mean drawn values,}
  \CommentTok{\# with one draw per raw}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ c2, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the mean of /k/ and /t/, for each draw}
    \AttributeTok{c2tk =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{c}\NormalTok{(c2k, c2t)),}
    \CommentTok{\# calculate the difference of /p/ and /t, k/}
    \AttributeTok{c2p\_tk\_diff =}\NormalTok{ c2p }\SpecialCharTok{{-}}\NormalTok{ c2tk}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrrrr@{}}

\caption{\label{tbl-voi-pbd}First 10 rows of expected difference of
voicing proportion between /t, k/ and /p/.}

\tabularnewline

\toprule\noalign{}
.draw & c2k & c2p & c2t & c2tk & c2p\_tk\_diff \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & -0.9918200 & -0.7450832 & -0.9166482 & -0.9533426 & 0.2082595 \\
2 & -1.0240673 & -0.7904348 & -1.0322325 & -0.9533426 & 0.1629079 \\
3 & -0.8954946 & -0.6873228 & -0.8904412 & -0.9533426 & 0.2660199 \\
4 & -0.9442361 & -0.8135191 & -0.9095122 & -0.9533426 & 0.1398235 \\
5 & -0.9278262 & -0.7908457 & -0.9200903 & -0.9533426 & 0.1624970 \\
6 & -0.9825895 & -0.7283054 & -0.9262196 & -0.9533426 & 0.2250372 \\
7 & -1.0482359 & -0.8731592 & -1.0841208 & -0.9533426 & 0.0801835 \\
8 & -1.0613713 & -0.8751219 & -1.0741555 & -0.9533426 & 0.0782208 \\
9 & -1.0622522 & -0.9739388 & -1.0481834 & -0.9533426 & -0.0205962 \\
10 & -0.9293998 & -0.8287910 & -0.9526105 & -0.9533426 & 0.1245517 \\

\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.9}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-diff-p-tk-1.pdf}}

}

\caption{\label{fig-diff-p-tk}Posterior distribution of the difference
between the expected voicing proportion of /t, k/ vs /p/, averaged
across vowels.}

\end{figure}%

Once we have the posterior difference, we can obtain CrIs of the
difference using \texttt{quantile2()} from the posterior package
\citep{burkner2024}. Beware that the values of the difference are in
log-odds! We can convert these into odd-ratios with \texttt{exp()}. The
output of the code below is shown in Table~\ref{tbl-voi-bm-quant}.
Odd-ratios indicate the ratio of the difference between A and B, so that
1 means no difference, values greater than 1 indicate an increase in A
relative to B and values lower than 1 indicate a decrease in A relative
to B. Odd-ratios are useful when looking at differences that are in
log-odds because while the relative magnitude of the difference in
proportion between two groups is the same independent of the baseline
proportion, the \emph{absolute} magnitude of the difference depends on
the baseline value. For example, an odd-ratio difference of 1.25 would
correspond to a proportion increase of about 3 percentage points if the
baseline proportion is 0.62 but it would correspond to a proportion
increase of 5 percentage points if the baseline proportion is 0.73 and 1
percentage point if the baseline is 0.99. Of course, in real research
contexts it is still useful to think about absolute magnitudes and their
relevance from a conceptual and methodological perspective. In this
tutorial we just focus on odd-ratios for simplicity.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(posterior)}

\NormalTok{voi\_bm\_quant }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{c2p\_tk\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# round to 2 digits}
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrr@{}}

\caption{\label{tbl-voi-bm-quant}Upper and lower limits of 90, 80 and
60\% Credible Intervals of the difference ratio of voicing proportion in
/t, k/ vs /p/.}

\tabularnewline

\toprule\noalign{}
q90 & q80 & q60 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1.07 & 1.10 & 1.14 \\
1.38 & 1.34 & 1.30 \\

\end{longtable}

Based on the model and data, there is a 90\% probability that the
voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38\% increase)
than in /t, k/. At 80\% confidence, the change ratio is 1.10-1.34 (or
10-34\% increase) while at 60\% confidence is 1.14-1.30 (14-30\%
increase). In other words we can be quite confident that the voicing
proportion in /p/ is longer than in /t, k/ and that the increase is less
than 35\%. The brms package comes with a convenient function,
\texttt{conditional\_effects()}, to plot posterior means and CrIs based
on predictors in the model. In Figure~\ref{fig-voi-prop-bm-cond}, we
plot the predicted proportion of voicing by consonant and vowel.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{conditional\_effects}\NormalTok{(voi\_prop\_bm, }\StringTok{"c2:vowel"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-cond-1.pdf}}

}

\caption{\label{fig-voi-prop-bm-cond}Expected voicing proportion by
consonant and vowel with 95\% Credible Intervals.}

\end{figure}%

Finally, the package marginaleffects \citep{arelbundock2025} has two
other convenience functions that return CrIs of comparisons across
predictor levels (\texttt{avg\_comparisons()}) and CrIs of posterior
predictions across predictor levels (\texttt{avg\_predictions}). Both
functions take the model and a list of variables as arguments.
Optionally, you can set the confidence level with the
\texttt{conf\_level} argument (here, we set it to 0.8, or 80\%).
\texttt{avg\_comparisons()} allows the user to specify the type of
comparison for each variable: here we set \texttt{"pairwise"} to
calculate pairwise comparisons of each level of \texttt{c2}. The output
of the two functions is shows in Table~\ref{tbl-marginal-1} and
Table~\ref{tbl-marginal-2}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(marginaleffects)}

\FunctionTok{avg\_comparisons}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \FunctionTok{list}\NormalTok{(}\AttributeTok{c2 =} \StringTok{"pairwise"}\NormalTok{), }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{type =} \StringTok{"link"}\NormalTok{)}
\FunctionTok{avg\_predictions}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \StringTok{"vowel"}\NormalTok{, }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}llrrr@{}}

\caption{\label{tbl-marginal-1}Difference in expected voicing proportion
by consonant, averaged across vowels, with 80\% CrIs.}

\tabularnewline

\toprule\noalign{}
term & contrast & estimate & conf.low & conf.high \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
c2 & mean(t) - mean(k) & 0.035 & -0.004 & 0.074 \\
c2 & mean(p) - mean(k) & 0.218 & 0.178 & 0.256 \\
c2 & mean(p) - mean(t) & 0.183 & 0.145 & 0.219 \\

\end{longtable}

\begin{longtable}[]{@{}lrrr@{}}

\caption{\label{tbl-marginal-2}Expected voicing proportion by vowel,
averaged across vowels, with 80\% CrIs.}

\tabularnewline

\toprule\noalign{}
vowel & estimate & conf.low & conf.high \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
i & 0.331 & 0.325 & 0.338 \\
e & 0.300 & 0.293 & 0.307 \\
a & 0.245 & 0.238 & 0.252 \\
o & 0.305 & 0.298 & 0.312 \\
u & 0.348 & 0.341 & 0.354 \\

\end{longtable}

\section{Case study 2: coarticulatory vowel
nasalisation}\label{sec-case-2}

For the second case study we will use data from \citet{carignan2021}.
The study looked at properties of nasality in German VNC sequences.
Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced
/d/) on the proportion of nasalisation within the vowel in the VNC
sequence. Previous work on coarticulatory nasalisation in English has
suggested that vowels followed by an NC sequence where C is voiceless
(NT) should show earlier coarticulatory nasalisation than vowels
followed by an NC sequence where C is voiced (ND). This pattern has been
suggested to be driven by the articulatory and acoustic incompatibility
of voicelessness and nasalisation, by which the velum opening gesture of
the nasal consonant is pushed away (i.e.~earlier) when the consonant
following the nasal is voiceless
\citep{fowler2005, beddor2009, cho2017, carignan2015}. Everything else
being equal, a greater proportion of vowel nasalisation (from the
perspective of time) should be found in vowels followed by NT than in
vowels followed by ND.

We will model the proportion of coarticulatory nasalisation in the
German short vowels /i, e, a, o, u/ when followed by /nt/ or /nd/, using
a Bayesian beta regression model. The proportion was calculated as the
proportion of the nasal interval to the duration of the vowel. The nasal
interval was defined thus: the interval between the time of peak
velocity of velum opening to the offset of the vowel. We will use the
results of the regression model to answer the following questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Is the nasalisation proportion, on average across vowels, greater in
  voiceless NC sequences?
\item
  Is there individual speaker variation?
\end{enumerate}

Table~\ref{tbl-nasal} shows the first ten rows of the data. The data
contains the following columns:

\begin{itemize}
\item
  \texttt{speaker} indicates the speaker ID.
\item
  \texttt{label} is the word label as given in the original data.
\item
  \texttt{vowel} is the target vowel in the VNC sequence.
\item
  \texttt{NC} is the NC sequence (voiceless /nt/ or voiced /nd/).
\item
  \texttt{voicing} indicates the voicing of C.
\item
  \texttt{nas\_prop} is the proportion of coarticulatory nasalisation of
  the vowel.
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lllllr@{}}

\caption{\label{tbl-nasal}First ten rows of the nasal proportion data.}

\tabularnewline

\toprule\noalign{}
speaker & label & vowel & NC & voicing & nas\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
S03 & b\_\emph{U\_nt}@\_\_\_N\_B17/s & u & nt & voiceless & 0.3668820 \\
S03 & b\_\emph{a\_nd}@\_\_\_N\_B19/s & a & nd & voiced & 0.1954858 \\
S03 & b\_\emph{a\_nt}@\_\_\_N\_B15/s & a & nt & voiceless & 0.2786485 \\
S03 & f\_\emph{I\_nt}@\_\_\_N\_B05/s & i & nt & voiceless & 0.7642259 \\
S03 & l\_\emph{I\_nd}@\_\_\_N\_B06/s & i & nd & voiced & 0.0052949 \\
S03 & p\_\emph{E\_nt}\_\_\_\_N\_B09/s & e & nt & voiceless &
0.3347331 \\
S03 & r\_\emph{a\_nt}@\_\_\_N\_B06/s & a & nt & voiceless & 0.2431760 \\
S03 & v\_\emph{I\_nd}@\_\_\_N\_B07/s & i & nd & voiced & 0.0247572 \\
S03 & v\_\_I\_nt\_6\_\_\_N\_B15/s & i & nt & voiceless & 0.1350081 \\
S03 & z\_\emph{E\_nd}@\_\_\_N\_B17/s & e & nd & voiced & 0.5378522 \\

\end{longtable}

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nasal-1.pdf}}

}

\caption{\label{fig-nasal}Proportion of coarticulatory nasalisation
during the vowel in VNC sequences in German, depending on C voicing.}

\end{figure}%

Figure~\ref{fig-nasal} shows the proportion of coarticulatory
nasalisation in vowels followed by /nd/ (voiced) vs /nt/ (voiceless)
sequences, for the short vowels /i, e, a, o, u/. We can see a pattern of
higher nasalisation proportion in vowels followed by /nt/, at least in
the vowels /a, i, o/. For /e, u/, the distribution of nasalisation
proportion seems to be similar between the voiced and voiceless
contexts.

Now onto modelling with a beta regression. Note that a full appropriate
model would include further predictors (both constant and varying), but
for simplicity here we include only the following predictors: voicing
(voiced /nd/ vs voiceless /nt/), vowel (/i, e, a, o, u/), including an
interaction between them. As varying terms, we include a varying
intercept by speaker and a by-speaker varying slope for voicing and
vowel in interaction. As with the model from the first case study,
voicing and vowel are coded using indexing, by suppressing the intercept
with \texttt{0\ +}. Here's the code of the model.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{  nas\_prop }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{|}\NormalTok{ speaker),}
  \AttributeTok{data =}\NormalTok{ nasal,}
  \AttributeTok{family =}\NormalTok{ Beta,}
  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
  \AttributeTok{file =} \StringTok{"data/cache/nas\_prop\_bm"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Let's inspect the output of \texttt{fixef()}, reported in
Table~\ref{tbl-nas-prop-fixef}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-nas-prop-fixef}Regression coefficients of a beta
regression of nasalisation proportion (\texttt{nas\_prop\_bm}).}

\tabularnewline

\toprule\noalign{}
& Estimate & Est.Error & Q10 & Q90 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
voicingvoiced:vowela & -0.61 & 0.12 & -0.77 & -0.46 \\
voicingvoiceless:vowela & -0.08 & 0.09 & -0.19 & 0.04 \\
voicingvoiced:vowele & -0.19 & 0.16 & -0.38 & 0.01 \\
voicingvoiceless:vowele & -0.25 & 0.10 & -0.38 & -0.12 \\
voicingvoiced:voweli & -0.74 & 0.19 & -0.98 & -0.51 \\
voicingvoiceless:voweli & -0.12 & 0.18 & -0.35 & 0.10 \\
voicingvoiced:vowelo & -0.34 & 0.16 & -0.54 & -0.14 \\
voicingvoiceless:vowelo & 0.25 & 0.15 & 0.05 & 0.44 \\
voicingvoiced:vowelu & -0.71 & 0.18 & -0.94 & -0.49 \\
voicingvoiceless:vowelu & -0.58 & 0.16 & -0.78 & -0.38 \\

\end{longtable}

Negative log-odds indicate a proportion that is smaller than 50\%, while
positive log-odds a proportion that is greater than 50\%. Generally, the
expected log-odd predictions in Table~\ref{tbl-nas-prop-fixef} are
negative, indicating an overall tendency for the nasalisation to take
less than 50\% of the duration of the vowel. Moreover, the predictions
are higher for voiceless NC sequences than for voiced NC sequences,
indicating a greater proportion of nasalisation in the former. However
there is vowel-specific variation, and there doesn't seem to be much of
a difference in nasalisation proportion in /e/ and /u/.
Figure~\ref{fig-nas-prop} shows the expected predictions with
\texttt{conditional\_effects()}, which should make the interpretation of
the predictions from Table~\ref{tbl-nas-prop-fixef} clearer.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{conditional\_effects}\NormalTok{(nas\_prop\_bm, }\StringTok{"vowel:voicing"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-1.pdf}}

}

\caption{\label{fig-nas-prop}Expected nasalisation proportion by vowel
and voicing, with 95\% CrIs.}

\end{figure}%

Now that we fitted the model we can use the posterior draws to answer
the two research questions (repeated from above):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Is the nasalisation proportion, on average across vowels, greater in
  voiceless NC sequences?
\item
  Is there individual speaker variation?
\end{enumerate}

To answer question 1, we can calculate the average difference in
nasalisation proportion by first calculating the average nasalisation
across all vowels for voiced and voiceless sequences (see
Table~\ref{tbl-nas-draws} for the output of this step) and then take the
difference of those, similarly to what we have done in the case study in
Section~\ref{sec-case-1}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{nas\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(nas\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}

\NormalTok{nas\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot vowel:c2 columns}
  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiced:vowela}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiceless:vowelu}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrrlllr@{}}

\caption{\label{tbl-nas-draws}First ten rows of posterior draws from
\texttt{nas\_prop\_bm} in long format.}

\tabularnewline

\toprule\noalign{}
.chain & .iteration & .draw & type & voicing & vowel & value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 1 & 1 & b & voicingvoiced & vowela & -0.4632870 \\
1 & 1 & 1 & b & voicingvoiceless & vowela & -0.0943699 \\
1 & 1 & 1 & b & voicingvoiced & vowele & -0.1546258 \\
1 & 1 & 1 & b & voicingvoiceless & vowele & -0.2080810 \\
1 & 1 & 1 & b & voicingvoiced & voweli & -0.4810383 \\
1 & 1 & 1 & b & voicingvoiceless & voweli & -0.0529377 \\
1 & 1 & 1 & b & voicingvoiced & vowelo & -0.3096219 \\
1 & 1 & 1 & b & voicingvoiceless & vowelo & 0.3212456 \\
1 & 1 & 1 & b & voicingvoiced & vowelu & -0.6198312 \\
1 & 1 & 1 & b & voicingvoiceless & vowelu & -0.3680900 \\

\end{longtable}

Now let's calculate the mean nasalisation proportion within each draw by
voicing, and plot the resulting posterior distributions. Note that, as
discussed in Section~\ref{sec-case-1}, when working with log-odds it is
important to first do all necessary calculations in log-odds (here
calculate the mean log-odds across vowels) and \emph{then} transform the
calculated estimands to proportions/probabilities. The output of the
following code is show in Table~\ref{tbl-nas-long-voi} and the density
plot of the calculated draws is in
Figure~\ref{fig-nas-prop-bm-draws-long-voicing-plot}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and voicing ensures that averaging applies only within draw and voicing}
  \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{summarise}\NormalTok{(}
    \CommentTok{\# calculate the mean value within draw/voicing in log{-}odds}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value),}
    \CommentTok{\# we can now transform log{-}odds to proportion with plogis()}
    \AttributeTok{value\_mean\_prop =} \FunctionTok{plogis}\NormalTok{(value\_mean),}
    \AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rlrr@{}}

\caption{\label{tbl-nas-long-voi}Mean nasalisation proportion by draw
and voicing, averaged across vowels (first 10 rows).}

\tabularnewline

\toprule\noalign{}
.draw & voicing & value\_mean & value\_mean\_prop \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & voicingvoiced & -0.4056808 & 0.3999482 \\
1 & voicingvoiceless & -0.0804466 & 0.4798992 \\
2 & voicingvoiced & -0.3905744 & 0.4035790 \\
2 & voicingvoiceless & -0.1227124 & 0.4693603 \\
3 & voicingvoiced & -0.4436082 & 0.3908816 \\
3 & voicingvoiceless & -0.1320190 & 0.4670431 \\
4 & voicingvoiced & -0.4385607 & 0.3920840 \\
4 & voicingvoiceless & -0.0981768 & 0.4754755 \\
5 & voicingvoiced & -0.3827134 & 0.4054726 \\
5 & voicingvoiceless & -0.2657250 & 0.4339569 \\

\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(value\_mean\_prop, voicing)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-bm-draws-long-voicing-plot-1.pdf}}

}

\caption{\label{fig-nas-prop-bm-draws-long-voicing-plot}Expected
nasalisation proportion by voicing, averaged across vowels.}

\end{figure}%

The plot suggests an overall greater nasalisation proportion in
voiceless NC sequences. Let's quantify how greater as we did in
Section~\ref{sec-case-1}. We will use odd-ratios in this context as
well, i.e.~we will convert log-odds to odd-ratios using the
\texttt{exp()} (exponential) function (and as before we first calculate
the difference and then exponentiate the resulting values, after which
we can take summary measures, like means and quantile-based measures
such as CrIs). Table~\ref{tbl-nas-prop-bm-diff-quant} show the 90, 80
and 60\% CrIs of the difference ratio of nasalisation proportion in
voiceless vs voiced NC sequences.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
  \CommentTok{\# with one draw per raw. we need to drop the value\_mean\_prop col}
  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{value\_mean\_prop) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the difference of voiceless and voiced in log{-}odds}
    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced,}
    \CommentTok{\# now transform with exp() to get the ratio difference}
    \AttributeTok{voicing\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(voicing\_diff)}
\NormalTok{  )}

\NormalTok{nas\_prop\_bm\_diff\_quant }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}rrr@{}}

\caption{\label{tbl-nas-prop-bm-diff-quant}Upper and lower limits of 90,
80 and 60\% Credible Intervals of the difference ratio of nasalisation
proportion in vowels followed by voiced vs voiceless consonants.}

\tabularnewline

\toprule\noalign{}
q90 & q80 & q60 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1.23 & 1.27 & 1.33 \\
1.69 & 1.63 & 1.56 \\

\end{longtable}

The CrIs of the ratio difference in nasalisation proportion in voiceless
vs voiced NC sequences suggest an increase of nasalisation in the
voiceless NC sequences, with a 90\% probability that the increase is
between 23\% and 69\% of the proportion in voiced NC sequences.

Moving onto question 2: is there individual speaker variation? While in
the recent tradition of linguistic research to use varying (aka random)
effects to \emph{control} for differences across participants, it has
been proposed to use the estimates of the varying coefficients to
investigate individual differences \citep{tamminga2016}. We will use the
\texttt{spread\_draws()} function from tidybayes \citep{kay2019} to
extract the draws of the varying terms (in brms these are the
coefficients that start with \texttt{r\_}). There is quite a few steps
of processing to get from the raw draws to the estimand we need: while
we have commented the following code, we encourage readers to test each
line sequentially and inspect the intermediate output to fully
understand the process. We assume that readers are familiar enough with
models with varying terms (aka random effects, mixed-effects models).
What readers should note is that to obtain the expected predictions of
nasalisation proportion for each speaker, the constant terms and the
varying terms should be added (since the varying terms indicate the
deviation of each speaker from the overall estimate).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidybayes)}

\NormalTok{nas\_prop\_r }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# extract draws from model, only \textasciigrave{}r\_speaker\textasciigrave{} varying terms}
  \FunctionTok{spread\_draws}\NormalTok{(r\_speaker[speaker,voicingvowel]) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# separate the column voicingvowel to two columns}
  \FunctionTok{separate}\NormalTok{(voicingvowel, }\FunctionTok{c}\NormalTok{(}\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# join the draws with the \textasciigrave{}b\_\textasciigrave{} terms}
  \FunctionTok{left\_join}\NormalTok{(}\AttributeTok{y =}\NormalTok{ nas\_prop\_bm\_draws\_long) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# get the expected log{-}odd value of each speaker, in each draw}
  \CommentTok{\# this is the sum of the \textasciigrave{}value\textasciigrave{} from the b\_ terms and the value from the}
  \CommentTok{\# r\_speaker term.}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{r\_speaker\_value =}\NormalTok{ value }\SpecialCharTok{+}\NormalTok{ r\_speaker) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# group the data for summarise}
  \FunctionTok{group\_by}\NormalTok{(.draw, speaker, voicing) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# get mean expected log{-}odds by draw, speaker and voicing (averaging across vowel)}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{r\_speaker\_value\_mean =} \FunctionTok{mean}\NormalTok{(r\_speaker\_value)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# make the data wider: two columns, one for voiced and one for voiceless}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ r\_speaker\_value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# finally, calculate the difference in expected log{-}odds of voiceless and voiced}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced)}
\end{Highlighting}
\end{Shaded}

Figure~\ref{fig-nas-prop-r} plots the posterior distributions of the
expected log-odd difference of coarticulatory nasalisation in voiceless
vs voiced NC sequences (\emph{x}-axis), for each speaker in the data
(\emph{y}-axis), as predicted by the model. The red solid vertical line
indicates the constant (overall) expected log-odd difference based on
the (constant) \texttt{b\_} terms. The black dashed vertical line marks
log-difference 0 (i.e., no difference in proportion of nasalisation
between voiceless and voiced NC).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_r }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(voicing\_diff, }\FunctionTok{reorder}\NormalTok{(speaker, voicing\_diff))) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \FunctionTok{mean}\NormalTok{(nas\_prop\_bm\_diff}\SpecialCharTok{$}\NormalTok{voicing\_diff), }\AttributeTok{colour =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{lims}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\FloatTok{1.5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-r-1.pdf}}

}

\caption{\label{fig-nas-prop-r}Expected log-odd difference of
coarticulatory nasalisation in voiceless vs voiced NC sequences for each
speaker.}

\end{figure}%

There is a lot of uncertainty within and between speakers: while the
distributions of most speakers are located in the positive range, some
expected distributions (see last 5 speakers at the bottom of figure) do
substantially span both negative and positive values. In other words,
while most speakers are more likely to have a larger nasalisation
proportion in voiceless NC sequences, a few might in fact have the
opposite pattern. Even among those speakers that do have a more robust
positive difference, there is a lot of uncertainty as to the magnitude
of the difference.

\section{Conclusion}\label{conclusion}

In this tutorial we have presented two case studies where the outcome
variable is a beta variable, i.e.~a numeric continuous variable bounded
between 0 and 1. We have shown how to fit Bayesian beta regressions to
coarticulatory progressive voicing in voiceless stops in Italian in
Section~\ref{sec-case-1} and coarticulatory regressive nasalisation in
vowels in German in Section~\ref{sec-case-2}. While this tutorial alone
will not be sufficient to be able to independently run full Bayesian
analyses, it will serve as a first introduction for readers to jump
start their learning.

  \bibliography{linguistics.bib,statistics.bib}

\end{document}