### 2025-02-10 README.qmd
render pdf and html

---
title: "Bayesian beta regressions with brms in R: A tutorial for phoneticians"
format: md
---

This is the GitHub repository of the paper *Bayesian beta regressions with brms in R: A tutorial for phoneticians* by Stefano Coretta and Paul Bürkner.

The HTML version of the paper can be found [here](manuscript/manuscript.html).