### 2025-02-10 README.qmd
render all

---
title: "Bayesian beta regressions with brms in R: A tutorial for phoneticians"
format: md
---

This is the GitHub repository of the paper *Bayesian beta regressions with brms in R: A tutorial for phoneticians* by Stefano Coretta and Paul Bürkner.

The code of the manuscript is in `manuscript/manuscript.qmd`. The `code` folder contains code with data checks and plotting.

The rendered HTML version of the paper and the data checks can be found on the [repository website](https://stefanocoretta.github.io/beta-phon/):

- [manuscript](manuscript/manuscript.html).
- [data checks](code/01_data_checks.html).